Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spread Across 400\,pc

### 2021-03-09 cluster_paper/ms.tex
propagate better vtang subtraction

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 300\,pc
  %An Open Cluster With a 300\,pc Halo
  An Open Cluster Spread Across 500\,pc
  %NGC\,2516  300\,pc
}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  500\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants, and
  expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

\section{A 300\,pc Halo around a Core?}
\label{sec:gaia}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf NGC\,2516 in the galactic context.} 
  Points denote halo members with $\pi/\sigma_\pi>20$.
  In our coordinate system, the Sun is at
  $\{X,Y,Z\} = \{ -8122, 0, +20.8\}\,{\rm pc}$.
  The gray line traces the past and future cluster orbit.  In
  different projections, the line spans different time intervals: the
  longest time window is visible in the $(Z,X)$ projection.
  The large blue arrows denote the median cluster velocity after
  subtracting the local standard of rest: 
  $\{v_{\rm X}, v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.
  Their sizes are internally proportionate.
  By comparison, $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
  \label{fig:XYZ}
	}
\end{figure*}

@@

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study in the space of positions, proper motions, and radial velocity
when available.  The \citetalias{cantatgaudin_gaia_2018} members are
all within a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members (the ``halo'') span tens of degrees. 
Inverting the parallax shows that members have been reported from
line-of-sight distances ranging from $300$ to $600\,$pc.
Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
$\mu_\delta$ projection were reported almost entirely by
\citet{meingast_2021}.

@@

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following ``cube'' in right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample, which led to
the omission of 1.1\% of the stars drawn from the volume noted above.

The discrepancy between our canonical understanding of open clusters
and the possibility of a diffuse population existing around \cn\
prompted our initial interest in this analysis.  What is the true
structure of \cn?  Are the core and halo truly coeval?

% FIXME FIXME FIXME
{\bf
To at least visualize the thing better...
}
Figure~\ref{fig:XYZ}.

@@

\begin{figure}[t]
	\begin{center}
			\includegraphics[width=0.49\textwidth]{f8a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Kinematic members of NGC\,2516, cross-matched against
  rotational members.} The projections are in galactic position
  coordinates and 2d-tangential velocity space.
  \label{fig:physical_x_rotn}
	}
\end{figure}

\begin{figure}[t]
	\begin{center}
			\includegraphics[width=0.49\textwidth]{f8b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
  versus 3d separation from the core and 2d velocity difference.} Bins
  have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in velocity.
  {\bf Uncertainties in the fraction are taken as the quadrature sum
  of the rel unc of the numerator and denominator (FIXME)}.  Bonafide
  members appear to exist hundreds of parsecs from the core, and up to
  $\approx$10\kms in tangential velocity separation.
  \label{fig:hist_physical_x_rotn}
	}
\end{figure}

How far away from the core, in position and velocity space, does the
halo extend?  We can explore this by crossmatching the stars with
detected rotation periods against our original target list of
\nkinematic Gaia DR2 kinematic candidate members.  For a fair
comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
which our TESS pipeline successfully made detrended light curves.  In
other words---the base sample is the stars for which we could have
plausibly detected a rotation period.  Faint stars with $Rp>16$ were
beyond our light curve selection limit, and crowded stars ({\it e.g.,}
near saturated stars in the cluster center) are similarly omitted.  We
selected the color limit above based on Figure~\ref{fig:rot}, since
the density of period detections seems to decrease once $(Bp-Rp)_0
\gtrsim 1.25$, {\it i.e.}, for spectral types later than $\approx$K4V.
Our expectation is that completion effects become important for
fainter stars.
%
% At the distance of NGC\,2516, a K4V star has $T\approx 14.5$,
% for which our reductions of the TESS images yield a precision of
% $\approx 7$\,mmag per half-hour exposure \citep{bouma_wasp4b_2019}.
% {\bf 
% The AMPLITUDES of the rotation signals are like 10-100 mmag (CITE
% REBULL 2020, FIGURE 10).
% }
%
%G=18 is like Abs G 9.8, Bp-Rp not corrected of 2.7. 
% (Bp-Rp )_0 of 1.25 is (Bp-Rp) of 1.38... or Abs G of ~= 6.8.
% or G-Rp ~= 0.7
% So, G~=15... meaning Rp~=14.3. 
%
% IN POSITION
% Bin width: 25
% Rotators: [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0
% 0   0   0 0   0]
% Comparison: [495 116  65  46  28  20  22  12  11   8   5   1   0   1
% 0   2   0   0 0   0]
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023
%

The results are shown using physical coordinates in
Figure~\ref{fig:physical_x_rotn}.  Appendix~\ref{app:gaia6d_x_rotn}
includes a supplemental visualization in the non-transformed space of
Gaia measurements.  To compute the physical positions, we transformed
from $(\alpha, \delta, \pi)$ to galactocentric ($X,Y,Z$) assuming the
\texttt{astropy v4.0} coordinate standard \citep{astropy_2018}.  We
then converted the on-sky proper motions to physical units by dividing
by the distance of each star, taken as the inverse parallax.  The
inverse parallax approximation does not introduce significant biases
for this sample: the mean $\pi/\sigma_\pi$ and its standard deviation
are 122 and 40 respectively; the minimum is 19.5.
Qualitatively, the fraction of kinematic members with TESS-detected
rotation periods seems like it may decrease with increasing distance
from the halo core---both in position and in velocity space.

To quantify this, we computed the separation of the positions and 2D
velocities from the median core member, selected by taking the median
across \citetalias{cantatgaudin_gaia_2018} members with membership
probabilities exceeding 70\%.  The result is shown in
Figure~\ref{fig:hist_physical_x_rotn}.  The halo extends to
separations of about 200\,pc in physical space from the cluster core.
This corresponds to total a length of $\approx$300-350\,pc, depending on
which members of the halo are chosen as the ``tips'' on either end.
Broadly speaking, this agrees with the results of
\citetalias{kounkel_untangling_2019}.

In tangential velocity space, the fraction of stars with rotation
period detections remains high out to roughly 10\kms.
\citet{meingast_2021} by comparison required a physically motivated
cut in tangential velocity space of 1.5\kms.

Of course, an important aspect of the tangential velocity cut needs to
be understood in light of 
PROJECTION EFFECTS.

... So uhhhhh yeah I need to compute that map.

While this 
improves purity, it is somewhat lacking in completeness...

% [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0   0]
% [495 116  65  46  28  20  22  12  11   8   5   1   0   1   0   2]
% [384 109  45  27  15   4   0   0   0   0   0   0   0   0   0   0   0   0
%    0   0]
% [533 153  67  40  23  10   3   1   2   0   0   0   0   0   0   0   0   0
%    0   0]

@@

\subsection{How did the halo form?}
Another way to frame this question: should we call these stars a tidal
tail?
% TODO: read Gagne+20 on this...
The classical explanation of such a structure (e.g., Krumholz+19) is
that stars escape out of L1 and L2, and then form leading and lagging
arms due to differential rotation in the galaxy.
The leading and lagging arms of NGC\,2516's halo qualitatively
match this picture, in that they are oriented roughly toward and away
the direction of galactic rotation (CITE, Meingast21, link to their
website).

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

@@

>>>>

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 300\,pc
  %An Open Cluster With a 300\,pc Halo
  An Open Cluster Spread Across 400\,pc
  %NGC\,2516  300\,pc
}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  $\approx$400\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants, and
  expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

\section{A Dispersed Halo and a Core?}
\label{sec:gaia}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.} 
  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
  cluster orbit.  In different projections, the line spans different
  time intervals: the longest time window is visible in the $(Z,X)$
  projection.  The large blue arrows denote the median cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.  The sizes of
  the blue arrows are internally proportionate.  By comparison,
  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
  \label{fig:XYZ}
	}
\end{figure*}

@@

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study in the space of observed positions, proper motions, and radial
velocity when available.  The \citetalias{cantatgaudin_gaia_2018}
members are all within a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members (the ``halo'') span tens of degrees.  Inverting the parallax
shows that members have been reported from line-of-sight distances
ranging from $300$ to $600\,$pc.  Stars in the upper right quadrant of
the $\mu_{\alpha'}$ {\it vs.} $\mu_\delta$ projection were reported
almost entirely by \citet{meingast_2021}.

@@

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following intervals of right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample, which led to
the omission of 1.1\% of the stars drawn from the volume noted above.

The discrepancy between our canonical understanding of open clusters
and the possibility of a diffuse population existing around \cn\
prompted our interest in understanding the reality and origin of the
diffuse population.  Figure~\ref{fig:XYZ} helps in visualizing the
proposed structure and kinematics of the halo.  We computed the
physical positions by transforming from $(\alpha, \delta, \pi)$ to
galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
standard \citep{astropy_2018}.  The direction of galactic rotation is
$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
median parameters for core members for which
\citetalias{cantatgaudin_gaia_2018} reported membership probabilities
exceeding 70\%.  We then integrated the orbit using \texttt{gala} and
the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.
The elongation of the cluster in both the $(X,Y)$ and $(Z,Y)$ planes
is correlated with the direction of the LSR-corrected median cluster
velocity.  A similar correlation was noted in Coma Ber by
\citet{tang_comaber_2019}.  Possible explanations are discussed in
Section~\ref{subsec:origin}.

@@

\begin{figure*}[t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f8a.pdf}
			\includegraphics[width=0.5\textwidth]{f8b.pdf}
		}
	%		\includegraphics[width=0.7\textwidth]{f8a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of NGC\,2516 exist hundreds of parsecs from the core,
  and up to $\approx$5\kms\ from the core in tangential velocity.} Stars
  that meet both kinematic and rotational membership criteria are shown in
  blue.  Stars for which rotation periods should have been detectable,
  but were not detected, are shown in orange.  Projections are shown in galactic coordinates and
  (projection-corrected) 2D-tangential velocity space.  The histogram
  on the right bins versus 3D separation from the core and
  2D velocity difference. Bins have width 25\,pc
  in position, and 1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties
  in the fraction are taken as the quadrature sum of the rel unc of
  the numerator and denominator (FIXME)}
  \label{fig:physical_x_rotn}
	}
\end{figure*}

% \begin{figure}[t]
% 	\begin{center}
% 			\includegraphics[width=0.49\textwidth]{f8b.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
%   \caption{ {\bf Histogram of Figure~\ref{fig:physical_x_rotn}, binned
%   versus 3d separation from the core and projection-corrected 2d velocity difference.} Bins
%   have width 25\,pc in position, and 1.0\,km\,s$^{-1}$ in velocity.
%   {\bf Uncertainties in the fraction are taken as the quadrature sum
%   of the rel unc of the numerator and denominator (FIXME)}.  Bonafide
%   members appear to exist hundreds of parsecs from the core, and up to
%   $\approx$5\kms in tangential velocity separation.
%   \label{fig:hist_physical_x_rotn}
% 	}
% \end{figure}

How far away from the core, in position and velocity space, does the
halo extend?  We can explore this by crossmatching the stars with
detected rotation periods against our original target list of
\nkinematic Gaia DR2 kinematic candidate members.  For a fair
comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
which our TESS pipeline successfully made detrended light curves.  In
other words---the base sample is the stars for which we could have
plausibly detected a rotation period.  Faint stars with $Rp>16$ were
beyond our light curve selection limit, and crowded stars ({\it e.g.,}
near saturated stars in the cluster center) are similarly omitted.  We
selected the color limit above based on Figure~\ref{fig:rot}, since
the density of period detections seems to decrease once $(Bp-Rp)_0
\gtrsim 1.25$, {\it i.e.}, for spectral types later than $\approx$K4V.
Our expectation is that completion effects become important for
fainter stars.

% IN POSITION
% Bin width: 25
% Rotators: [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0
% 0   0   0 0   0]
% Comparison: [495 116  65  46  28  20  22  12  11   8   5   1   0   1
% 0   2   0   0 0   0]
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023
%

The results are shown using physical coordinates in
Figure~\ref{fig:physical_x_rotn}.  Appendix~\ref{app:gaia6d_x_rotn}
includes a supplemental visualization in the observed space of Gaia
measurements.  The observed proper motions however suffer from
projection effects, discussed in Appendix~\ref{app:vproj}.  These
effects are important in the equatorial coordinate system due to the
proximity of some \cn\ members to the southern ecliptic pole.  For a
star at position $(\alpha, \delta)$, we correct for the projection
effect by comparing the observed proper motion $(\mu_{\alpha'},
\mu_\delta)$ to what the proper motion at the star's position would
have been if the star were comoving with the core of \cn.  This yields
a quantity we denote $\Delta \mu^{*}$, following
\citetalias{meingast_2021}.  We then convert these proper motion
differences to physical units by dividing by the distance of each
star, taken as the inverse parallax.  The inverse parallax
approximation does not introduce significant biases for this sample,
since the mean $\pi/\sigma_\pi$ and its standard deviation are 122 and
40 respectively; the minimum is 19.5.  An inspection of the results
suggests that the fraction of kinematic members with TESS-detected
rotation periods may decrease with increasing distance from the halo
core---both in position and in velocity space.

To quantify this, we compute the separation of the positions and 2D
velocities from the median \citetalias{cantatgaudin_gaia_2018} core
member.  The result is shown on the right side of
Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
about 200\,pc in physical space from the cluster core.  This
corresponds to total a length of $\approx$400\,pc, depending on which
members of the halo are chosen as the ``tips'' on either end.  This
agrees broadly speaking with the structure of the halo reported by
\citetalias{kounkel_untangling_2019}.

In projection-corrected tangential velocity space, the fraction of
stars with rotation period detections remains high out to roughly
5\kms.  \citet{meingast_2021} by comparison required a physically
motivated cut in tangential velocity space of 1.5\kms.  
Our results show that at the expense of higher field star
contamination rates, bonafide members can be identified even out to
higher velocity separations.
{\bf todo: anything else to mention? e.g., the actual numbers?}

@@

\subsection{How did the halo form?}
\label{subsec:origin}

Differential rotation in the galaxy is clearly important. Meingast+21
established this; in ten different cluster halos, nine showed leading
and trailing arms oriented in line with the differential rotation of
the Galaxy.
(See e.g., their Figure 13).

Another way to frame this question: should we call these stars a tidal
tail?
% TODO: read Gagne+20 on this...
The classical explanation of such a structure (e.g., Krumholz+19) is
that stars escape out of L1 and L2, and then form leading and lagging
arms due to differential rotation in the galaxy.
The leading and lagging arms of NGC\,2516's halo qualitatively
match this picture, in that they are oriented roughly toward and away
the direction of galactic rotation (CITE, Meingast21, link to their
website).

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

@@

\section{Projection effects for 2D velocities}
\label{app:vproj}

Figure~\ref{fig:vproj} shows why the correct approach for comparing
any star's proper motions against the cluster's mean depends on the
position of the star being considered.
